\normalfont \normalsize \textsc{Math 460} \\ [25pt] \horrule{0.5pt} \\[0.4cm] \huge Golay Codes\\ \horrule{2pt} \\[0.5cm]

### 2015-05-05 paper.tex
moving to the PC
first 2/3 written

\usepackage{tikz}
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[american]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm} % Math packages

@@

\section{History and Construction}
\begin{outline}[enumerate]
   \1 Constructions of the perfect Golay $[23,12,7]$ code $\mathcal{G}_{23}$ and extended (not perfect) $[24,12,8]$ Golay $[24,12,8]$ code $\mathcal{G}_{24}$.
       \2 As the $S(5,8,24)$ Steiner System [\cite{griess}]
         \3 Steiner Systems
           \begin{defn}{Steiner system}

            A Steiner system with parameters (a,b,c) is a family $S$ of subsets of some set $\Omega$ of c elements such that (i) $A \in S$ implies that $|A| = b$; (ii) for any a-set \textit{F} in $\Omega$, there is a unique $A \in S$ such that $\textit{F} \subseteq A$.
           \end{defn}
         \3 Fano plane as example of S(2,3,7). Seven lines partition the seven points, where each line contains exactly three points, and each pair of points uniquely designates a line.
\begin{tikzpicture}

@@

\end{tikzpicture}

      \2 Generator Matrix as [I:A] where A is the adjacency matrix of the icosahedron.
      \2 Perfect $\mathcal{G}_{23}$ created by removing the parity digit.
\end{outline}

\section{Important Properties}
\begin{outline}[enumerate]
  \1 Important properties of the Golay Codes
       \2 Octads, dodecads, and sextets
         \begin{defn}{Octads, dodecad, and sextets}

           For $S(5,8,24)$ we call elements of $S$ octads, a dodecad is a 12-set, a sextet is a 4-set (the sextets partition the space into six set).
         \end{defn}
      \2 Minimum weight of $\mathcal{G}_{24}$ is 8, other codeword weights are 12, 16, and 24. We can detect 7 errors and correct $3 = \lfloor \frac{8-1}{2} \rfloor$. This is where we see that $\mathcal{G}_{24}$ is not perfect because spheres of radius 4 centered at codewords intersect while spheres of radius 3 do not fill the space. 
      \2 Self-duality [\cite{ivanov}]
         \begin{defn}{dual code}

          For a code $\mathcal{C}$, the dual code $\mathcal{C}^*$ is the orthogonal complement of $\mathcal{C}$ with respect to the parity form. ie. $\mathcal{C}^{*} = \{A | A \in 2^X, |A \cap B|$ is even for all $B \in \mathcal{C} \}.$
         \end{defn}
         \begin{defn}{totally singular}

          We call a code $\mathcal{C}$ totally singular if $\mathcal{C} \subseteq \mathcal{C}^*$.
         \end{defn}
         \begin{lma}
          A code $\mathcal{C}$ of length $X$ is self-dual if it is totally singular with respect to the parity form and $dim(\mathcal{C}) = |X|/2$.
         \end{lma}
\end{outline}
\section{The Automorphism Group $M_{24}$}
\begin{outline}[enumerate]
   \1 Automorphism Group of $\mathcal{G}_{24}$.       
       \2 The hexacode and Miracle Octad Generator (MOG) are valuable tools when working with $\mathcal{G}_{24}$ because it gives us a method to quickly check whether a message is a valid codeword. It also becomes useful when investigating maximal subgroups of $Aut(\mathcal{G}_{24})$ [\cite{conway}].

       \begin{defn}{Hexacode}

        Consider $\mathbb{F}_4 = \{0,1,\alpha,\bar{\alpha}\}$ with the usual relations.
        Then hexacode $\mathcal{C}_6$ has a word $W(\phi) = ab|cd|ef$ for each quadratic function $\phi(x) = ax^2 + bx + c$ defined over $\mathbb{F}_4$. Where $c=\phi(0)$, $d=\phi(1)$, $e=\phi(\alpha)$, $f=\phi(\bar{\alpha})$.

        We can also consider $\mathcal{C}_6$ to be generated by 
        $$\alpha\bar{\alpha}|\alpha\bar{\alpha}|\alpha\bar{\alpha}$$
        $$\alpha\bar{\alpha}|\bar{\alpha}\alpha|\bar{\alpha}\alpha$$
        $$\bar{\alpha}\alpha|\alpha\bar{\alpha}|\bar{\alpha}\alpha$$
        $$\bar{\alpha}\alpha|\bar{\alpha}\alpha|\alpha\bar{\alpha}.$$
       \end{defn}
       \2 k-fold transitivity
          \begin{defn}
             Let G be a group acting on a set $\Omega$, and $k \leq |\Omega|$. We say that G is k-fold transitive on $\Omega$ if, given any two ordered k-tuples $(x_1, \ldots, x_k)$ and $(y_1, \ldots, y_k)$, there is an element $g \in G$ such that $x_ig = y_i$ for all $i \in {1, \ldots, k}$. If such a $g \in G$ is unique, then we call G sharply k-fold transitive on $\Omega$.
          \end{defn}
       \2 We will find that $Aut(\mathcal{G}_{24}) = M_{24}$ where $M_{24}$ is a 5-fold transitive, simple permutation group on 24 points. We can also show that $|M_{24}| = 2^{10} \times 3^3 \times 5 \times 7 \times 23 = 244,823,040$.

\end{outline}
\clearpage
\printbibliography
%----------------------------------------------------------------------------------------

>>>>

\usepackage{svg}
\usepackage{tikz}
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
%\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default

\usepackage[american]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage{bm}

@@

\section{History and Constructions}

The Golay code developed by Marcel Golay in 1949 has had significant impact on the advancement of coding theory, the classification of simple finite groups, and communication systems. Golay's 1949 article ``Notes on Digital Coding'' appeared in the June edition of the Proceedings of the IRE [\cite{golay}]. In ``Notes on Digital Coding'' Golay presented a generalization of the Hamming(7,4) code which he had read about in Shannon's 1948 paper. After generalizing the perfect, single error-correcting codes, Golay then justifies the existence of two codes which can correct multiple errors, the perfect binary Golay code $\mathcal{G}_{23}$, and the ternary Golay code. At the bottom of the article Golay supplied generator matrices for his codes, but lacking any explanation of their construction. Golay would provide explanation of their construction in 1954. Though brief and without significant explanation, Berlekamp ``the best single published page'' and Zaremba once told Golay that he had ``said more in half a page than others have said in twenty pages'' [\cite{thompson}].

The remainder of this section will concentrate on Golay's original construction of $\mathcal{G}_{23}$, some of the special properties that $\mathcal{G}_{23}$ and $\mathcal{G}_{24}$ statisfy. Then we can motivate the Golay Codes as steiner systems which we can show to be unique and equipped with known automorphism groups. These connections are the tools that allowed $\mathcal{G}_{24}$ to be used in the classification of finite simple groups and in particular aided in the discovery of Conway's sporadic groups $Co_1$,$Co_2$, and $Co_3$ [\cite{thompson}].

Golay first used Pascal's triangle to find 3 binomial coefficients which summed to a power of 2. The 23rd row of Pascal's triangle indicates that
$$\sum_{i=0}^3 {23 \choose i} = 1+23+253+1771 = 2048 = 2^{11}$$. This tell us that on the 23-cube, a sphere with Hamming radius atleast 3 contains a number of vertices which divides $2^{23}$ i.e. $V(S_{\geq 3}(x) )|2^{23}$ where $x$ is a vertex of the 23-cube. This implies that the minimum weight of $\mathcal{G}_{23}$ must be equal to $2*3+1=7$. Without justification Golay provided the generator matrix $\mathbf{A}$ for $\mathcal{G}_{23}$:

$$\textbf{A} = \begin{array}{ccccccccccccc}
& Y_1 & Y_2 & Y_3 & Y_4 & Y_5 & Y_6 & Y_7 & Y_8 & Y_9 & Y_{10} & Y_{11} & Y_{12} \\
   X_1 & 1 & 0 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 1 & 1 & 1\\
   X_2 & 1 & 0 & 1 & 0 & 1 & 1 & 0 & 1 & 1 & 0 & 0 & 1\\
   X_3 & 1 & 0 & 1 & 1 & 0 & 1 & 1 & 0 & 1 & 0 & 1 & 0\\
   X_4 & 1 & 0 & 1 & 1 & 1 & 0 & 1 & 1 & 0 & 1 & 0 & 0\\
   X_5 & 1 & 1 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 1 & 0 & 0\\
   X_6 & 1 & 1 & 0 & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 1\\
   X_7 & 1 & 1 & 0 & 1 & 1 & 0 & 0 & 1 & 1 & 0 & 1 & 0\\
   X_8 & 1 & 1 & 1 & 0 & 0 & 1 & 0 & 1 & 0 & 1 & 1 & 0\\
   X_9 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 0 & 0 & 0 & 1 & 1\\
X_{10} & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 1\\
X_{11} & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
\end{array}$$
Using this matrix we can map data vectors in $(\mathbb{Z}/2)^{12}$ to parity check words in $(\mathbb{Z}/2)^{11}$. Then for a data vector $\mathbf{v} \in (\mathbb{Z}/2)^{23}$ we can build the message $\mathbf{m} = (\mathbf{v},\mathbf{A}\mathbf{v})$. The parity check matrix is then $[\mathbf{A}:\mathbf{I}_{11}]$, and is a linear transformation $$[\mathbf{A}:\mathbf{I}_{11}]: (\mathbb{Z}/2)^{23} \longrightarrow (\mathbb{Z}/2)^{11},$$ where $\mathcal{G}_{23}=Ker([\mathbf{A}:\mathbf{I}_{11}])$. Thus to check to see if a received message $\mathbf{m}$ has an error we just check $\mathbf{Pm} \equiv 0 (mod 2)$.

The construction given by Golay for matrix $\textbf{A}$ depends upon a sketch of five lines.

\begin{figure}[!ht]
  \centering
  \def\svgwidth{150pt}
  \input{lines.pdf_tex}  
  \caption{Golay's lines}
  \label{fig:cstori}
\end{figure}

Then we label the minor matrix consisting of rows $X_1 \ldots X_{10}$ and columns $Y_2 \ldots Y_{12}$.

$$\textbf{A} = \begin{array}{ccccccccccccc}
&&       A  &  B  &  C  &  D  &  E &\sigma_1&\sigma_2&\sigma_3&\sigma_4&\sigma_5&\sigma_6\\
&&      Y_2 & Y_3 & Y_4 & Y_5 & Y_6 & Y_7 & Y_8 & Y_9 & Y_{10} & Y_{11} & Y_{12} \\
AB &    X_1 & 0 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 1 & 1 & 1\\
AC &    X_2 & 0 & 1 & 0 & 1 & 1 & 0 & 1 & 1 & 0 & 0 & 1\\
AD &    X_3 & 0 & 1 & 1 & 0 & 1 & 1 & 0 & 1 & 0 & 1 & 0\\
AE &    X_4 & 0 & 1 & 1 & 1 & 0 & 1 & 1 & 0 & 1 & 0 & 0\\
BC &    X_5 & 1 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 1 & 0 & 0\\
BD &    X_6 & 1 & 0 & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 1\\
BE &    X_7 & 1 & 0 & 1 & 1 & 0 & 0 & 1 & 1 & 0 & 1 & 0\\
CD &    X_8 & 1 & 1 & 0 & 0 & 1 & 0 & 1 & 0 & 1 & 1 & 0\\
CE &    X_9 & 1 & 1 & 0 & 1 & 0 & 1 & 0 & 0 & 0 & 1 & 1\\
DE & X_{10} & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 1\\
\end{array}$$
$$\sigma_1=(ABEDC), \sigma_2=(ABCED), \sigma_3=(ABDCE),$$
$$\sigma_4=(ACEBD), \sigma_5=(ACBDE), \sigma_6=(ADCBE)$$
Then for columns $Y_2 \ldots Y_6$ we place a 0 in the rows corresponding to intersections with the respective line, and 1 otherwise. In columns $Y_6 \ldots Y_{12}$ we place 0's in the rows where the corresponding transposition appears in the cycle, and ones elsewhere. The resulting matrix has rows and columns of minimum weight 6, so 
$$Y_1 = [ \begin{array}{ccccccccccc} 1& 1& 1& 1& 1& 1& 1& 1& 1& 1& 0\end{array}]^T$$
$$X_{12} = [\begin{array}{cccccccccccc} 0& 1& 1& 1& 1& 1& 1& 1& 1& 1\end{array}]$$ fixes the minimum weight to 7 [\cite{thompson}]. To construct $\mathcal{G}_{24}$ we append a 1 to each codeword of odd weight in $mathcal{G}_{23}$ and a zero to all codewords of even weight. This increases the minimum weight to 8, so while the code is no longer perfect, it can detect 7 errors rather than the 6 detectable by $\mathcal{G}_{23}$.

While the above recounts Golay's original construction of the generator matrix of $G_{23}$, history has demonstrated that the $G_{24}$ code is much more useful and more intuitive constructions of either code are available.

\subsection{As a Cyclic Code} (might cut this section)
We can construct $\mathcal{G}_{23}$ more easily by considering the factors of $$x^{23}-1=(x+1)(x^{11}+x^9+x^7+x^6+x^5+x+1)(x^{11}+x^{10}+x^6+x^5+x^4+x^2+1)$$ over $\mathbb{Z}/2$. If we consider $p(x)=x^{11}+x^9+x^7+x^6+x^5+x+1$ (though either degree 11 factor will generate the same code), then we know since $p|x^{23}-1$, that $<p(x)>$ is an ideal in the ring $\mathbb{Z}/2[x]$. Then $<p(x)>$ is closed under multiplication, in particular for any $g(x)\in<p(x)>$, we know $x*g(x)\in<p(x)>$. Further, in $\mathbb{Z}/2[x]$, multiplication by $x$ corresponds to the shift operator. The corresponding matrix for this code is 
$$\textbf{G} = \begin{array}{ccccccccccccc}
       p(x) & 1 & 0 & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 1 & 1\\
   x^1*p(x) & 1 & 1 & 0 & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 1\\
   x^2*p(x) & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0\\
   x^3*p(x) & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 1 & 1 & 0 & 0\\
   x^4*p(x) & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 1 & 1 & 0\\
   x^5*p(x) & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 1 & 1\\
   x^6*p(x) & 1 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 1\\
   x^7*p(x) & 1 & 1 & 0 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 1\\
   x^8*p(x) & 1 & 1 & 1 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 1 & 0\\
   x^9*p(x) & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 1\\
x^{10}*p(x) & 1 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 1 & 1 & 0 & 0\\
\end{array}$$

\subsection{As a Steiner System}
This is one of the most interesting constructions of the Golay code as it uses particular properties of the code to show that it satisfies the Assmus-Mattson Theorem which proves the existence of the Steiner system construction. Once we know that our construction is valid, we can use properties of Steiner systems to prove the uniqueness of the Golay code and to identify the automorphism group of the $\mathcal{G}_{24}$ code.

\begin{defn}
Consider integers $1 \leq t < k < v$ and $\lambda$. The design $t-(k,v,\lambda)$ is a pair of sets $(\mathfrak{X},\mathfrak{B})$ where $|\mathfrak{X}|=v$ and $\mathfrak{B}$ contains size $k$ subsets of $\mathfrak{X}$, called `blocks,' such that every size $t$ subset of $\mathfrak{X}$ is contained in exactly $\lambda$ blocks. In the particular case that $\lambda=1$, we call it a \textit{Steiner system} S(t,k,v).
\end{defn}

The Fano plane is the geometry associated with S(2,3,7). Seven lines partition the seven points, where each line contains exactly three points, and each pair of points uniquely designates a line.

\begin{figure}[hbp]
\centering
\begin{tikzpicture}[scale=2.5]

@@

\end{tikzpicture}
\caption{Fano Plane S(2,3,7)} \label{fig:fplane}
\end{figure}

In particular we want to use the Assmus-Mattson theorem to find the designs corrsponding to the Golay codes. Doing so will require several definitions.
\begin{defn}
Suppose $\mathcal{C}$ is a linear code in vectorspace $V$. We denote 
$$\mathcal{C}^* = \{ \bm{u} \in V | \bm{u} \cdot \bm{w} = 0 \; \forall \; \bm{w} \in V \}. $$
If $\mathcal{C}$ is an $[n,k]$ code, then $\mathcal{C}^*$ is an $[n,n-k]$ code.
Furthermore, if $\mathcal{C}\subseteq\mathcal{C}^*$ then we call $\mathcal{C}$ \textit{totally singular}.
\end{defn}
We note that $\mathcal{G}_{24}$ is totally singular, because the rows of its generator matrix are orthogonal.
More explicitely $\bm{r}_i \cdot \bm{r}_j$ = 0 for any two rows in the matrix $\bm{A}$ of the generator matrix $[\bm{I}_{12}:\bm{A}]$ above. This implies $\mathcal{G}_{24} \subseteq \mathcal{G}_{24}^*$.
\begin{defn}
If $\mathcal{C}$ is totally singular and $dim(\mathcal{C})=\frac{dim(V)}{2}$ then we can $\mathcal{C}$ is \textit{self-dual}.
\end{defn}
We just showed that $\mathcal{G}_{24}$ is totally singular, and we already know that $dim(\mathcal{G}_{24}) = 12 = \frac{24}/2$, so $\mathcal{G}_{24}$ is self-dual.
\begin{defn}[We might not need this definition]
A code $\mathcal{C}$ is \textit{double-even} if for every $x\in \mathcal{C}$, the weight of $c$ is divisible by 4.
\end{defn}
\begin{thm}
Let $\mathcal{C}$ be an $[n,k,d]$ binary code, and let $0<t<d$. Let $B_i$ be the number of vectors of weight $i$ in $\mathcal{C}^*$, and let
$$
s = |{i|B_i \not= 0 \; \& \; 0<i\leq n-t}|.
$$
If $s\leq d-t$, then the vectors of weight $d$ in $\mathcal{C}$ hold a t-design and the vectors of any weight $i\leq n-t$ and $B_i\not=0$ hold a t-design.
\end{thm}

As we've shown that $\mathcal{G}_{24}=\mathcal{G}_{24}^*$, and we know that we can correct 3 errors in any weight 8 word. We are motived to show that $\mathcal{G}_{24}$ contains a $5-(8,24,1)$ design. Since, all vectors in $\mathcal{G}_{24}$ are of weight 8, 12, or 16, we know that $s=3 \leq 8-5$. So the weight eight words contain the Steiner system S(5,8,24). Similar systems can be identified for code words of weight 12 and 16 [\cite{pless}].

\section{Uniqueness of $\mathcal{G}_{24}$}
This section will briefly cover the uniqueness proof of $\mathcal{G}_{24}$, these are important as they show that our three constructions actually generate the same code.
\section{The Automorphism Group of $\mathcal{G}_{24}$, $M_{24}$}
Knowing that our constructions are all equiavlent, we can leverage how $\mathcal{G}_{24}$ behaves under $M_{24}$. The Steiner system construction is important here because $M_{24}$ was first understood as the automorphism group of S(5,8,24).

\clearpage
\printbibliography
%----------------------------------------------------------------------------------------
